### 2018-12-10 writing/thesis/chapters/methods.tex
Fix limiter condition.
Fix some errors in description of dg problem.

% Variables and other equation stuff
\newcommand{\Q}{\bm{Q}}
\newcommand{\gradQ}{\gradient{\Q}}
\newcommand{\Qrho}{\rho}
\newcommand{\Qj}{\rho \bm{v}}
\newcommand{\Qv}{\bm{v}}
\newcommand{\QE}{\rho E}
\newcommand{\QZZ}{Z} % TODO: Name?
\newcommand{\QZ}{\rho \QZZ}
\newcommand{\potT}{\theta}
\newcommand{\backgroundPotT}{\overline{\theta}}
\newcommand{\pertubationPotT}{\theta'}
\newcommand{\stressT}{\bm{\sigma}}
\newcommand{\pressure}{p}
\newcommand{\maxConvEigen}{\vert \lambda_c^{\text{max}} \vert}
\newcommand{\maxViscEigen}{\vert \lambda_v^{\text{max}} \vert}
\newcommand{\Riemann}{\operatorname{Riemann}}

@@

\newcommand{\dsol}{\bm{u}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}
\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law} by a test function $\testfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
We arrive at the so called weak formulation of the \textsc{pde} 
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\testfunction{k} \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k} \left( \div{F(\Q, \gradQ} \right)
}}
=
\intdt{\intdcell{
    \testfunction{k} S(\Q, \bm{x}, t)
}}
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
We now replace the solution $\bm{Q}$ with so called spacetime-predictor $\stpredictor (\bm{x},t)$ into the weak form and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\intdcell{
  \testfunction{k} \sbasis{l}
}
\right)
(\bm{u^{n+1} - u^{n}})
&+
\left(\intdt{\intdcell{
      \testfunction{k} \Riemann(\stpredictor^-, \stpredictor^+) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\testfunction{k}} \cdot  \flux(\stpredictor)
}}\right)
&=
\left(\intdt{\intdcell{
      \testfunction{k} \source(\stpredictor)
}}\right)
\end{split}
\end{align}
The first term motivates an orthogonal basis as this leads to a diagonal mass matrix.

\sidetitle{Predictor}
To derive the predictor, we again take an approximation of our solution in a nodal basis, but now consider polynomials that are defined both in space and time
\begin{equation}
  \label{eq:cell-approx-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis{l}(\bm{x}, t).
\end{equation}
We now multiply the conservation law again by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k}{\bm{x}, t}
    \left(
      \divergence{\flux(\stpredictor)}
    \right)
}}
=
\intdt{\intdcell{
  \testfunction{k} \source(\stpredictor)
}}.
\end{equation}
Similar to the derivation of the corrector, we again integrate the first term by parts in time and the flux divergence in space.
This time we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
Note that this neglects the interaction with neighbouring cells; this is corrected in the corrector step.
\todo{Replace $\bm{x}$ with a macro some other letter. Is also coordinate!}

\begin{equation}
  \label{eq:space-time-predictor}
0
  + \ldots
\end{equation}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be solved in a fixed point iteration scheme.
For details and proof of convergence for the linear case, see (todo).
Write about initial guess (we use naive one anyway, right?)

\sidetitle{Boundary extrapolation and stuff}
Extrapolate unknowns to boundary, convert from our basis to basis on faces.
Then Riemann solver
and so on

@@

\sidetitle{Diffusive Flux stuff}
Numerical flux:
\begin{equation}
  \label{eq:rusanov-flux}
  \Riemann(q_h^-, \nabla q_h^-; g_h^+, \nabla q_h^+) \cdot \bm{n} =
  \frac{1}{2} \left(
    F(q_h^+, \nabla q_h^+) +
    F(q_h^-, \nabla q_h^-)
  \right) -
  \frac{1}{2} s_\text{max} (q_h^+ - q_h^-)
\end{equation}
with
\begin{equation}
  \label{eq:parabolic-penalty}
  s_\text{max}  = \max \left(
\vert \lambda_c(q_h^-) \vert, \, \vert \lambda_c(q_h^+)
\right) +
2 \eta \max \left(
\vert \lambda_v(q_h^-) \vert, \, \vert \lambda_v(q_h^+)
\right)
\end{equation}
and
\begin{equation}
  \eta = \frac{2N+1}{h \sqrt{\frac{1}{2} \pi}}
\end{equation}

Maximum timestep (\textsc{cfl})
\begin{equation}
 \Delta t \leq  \text{CFL} \, \frac{\alpha(N) \, h}{\maxConvEigen + 2 \maxViscEigen \frac{2N+1}{h}}
\end{equation}
with $N$ polynomial order and $h$ characteristic length scale of elements~\cite{dumbser2010arbitrary,gassner2008discontinuous}.
The constant $\alpha(N) \leq \frac{1}{2N+1}$ is obtained from von Neumann analysis on a simple model problem and depends on the approximation order~\cite{dumbser2008unified}.

\section{Reactive Compressible Navier Stokes Equations}
Fluid motion can be described by the compressible Navier Stokes equations.
We follow the description in~\cite{dumbser2010arbitrary} for the Navier Stokes part.
The coupling with the advection-diffusion-reaction equation follows~\cite{hidalgo2011ader}.

The vector of conserved quantities is given by
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right).
\end{equation}
where $\Qrho$ is the density, $\Qj$ is the momentum, $\QE$ the energy density and $\QZ$ is the mass fraction of the chemical reactant.

We split the flux into a hyperbolic part $\hyperFlux(\Q)$ and an viscous flux $\viscFlux(\gradQ)$
\begin{equation}
  \label{eq:flux}
  \flux(\Q, \gradQ) = \hyperFlux(\Q) + \viscFlux(\Q, \gradQ)
\end{equation}

@@

Complete \textsc{pde}:
\begin{equation}
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE\\
    \QZ
    \end{pmatrix}}_{\Q}
  + 
  \divergence{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}
 =
  \underbrace{
  \begin{pmatrix}
    S_{\Qrho\phantom{\Qrho}}\\
    S_{\Qj}\\
    S_{\QE}\\
    S_{\QZ}
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\quad
 \begin{array}{l}
 \text{mass cons.} \\
 \text{momentum cons.} \\
 \text{energy cons.} \\
 \text{cont.\ gas} 
\end{array}
\end{equation}

@@

A relatively recent way of dealing with this problem is the finite volume subcell limiter.
Our discussion follows the one in~\cite{dumbser2016simple}.
This limiting is an a posteriori limiting, which means that we first evaluate a timestep with an unlimited \textsc{ader-dg} method and then check, whether our solution is \enquote{correct}.
In our case, we check whether the solution is a valid floating point number.
We also check whether it is physically admissible.
In our case we use the criterion
\begin{equation}
  \label{eq:limiting-physical}
  \operatorname{is-admissible}(\Q) =
  \begin{cases}
    \text{true} & \Qrho > 0 \land \QE > 0 \\
    \text{false} & \text{otherwise}.
  \end{cases}
\end{equation}
\todo{Mention dmp if I end up using it.}

>>>>

% Variables and other equation stuff
\newcommand{\Q}{\bm{Q}}
\newcommand{\gradQ}{\gradient{\Q}}
\newcommand{\Qrho}{\rho}
\newcommand{\Qj}{\rho \bm{v}}
\newcommand{\Qv}{\bm{v}}
\newcommand{\QE}{\rho E}
\newcommand{\QZZ}{Z} % TODO: Name?
\newcommand{\QZ}{\rho \QZZ}
\newcommand{\potT}{\theta}
\newcommand{\backgroundPotT}{\overline{\theta}}
\newcommand{\pertubationPotT}{\theta'}
\newcommand{\stressT}{\bm{\sigma}}
\newcommand{\pressure}{p}
\newcommand{\maxConvEigen}[1][]{
  \vert%
  \lambda_c^{\text{max}}
  \notblank{#1}{\left(#1\right)}{}
  \vert%
}
\newcommand{\maxViscEigen}[1][]{
  \vert%
  \lambda_v^{\text{max}}
  \notblank{#1}{\left(#1\right)}{}
  \vert%
}
\newcommand{\Riemann}{\operatorname{Riemann}}

@@

\newcommand{\dsol}[1][h]{\bm{u}_{#1}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}
\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law} by a test function $\testfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
We arrive at the so called weak formulation of the \textsc{pde} 
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\testfunction{k} \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k} \left( \div{F(\Q, \gradQ} \right)
}}
=
\intdt{\intdcell{
    \testfunction{k} S(\Q, \bm{x}, t)
}}
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
We now replace the solution $\bm{Q}$ with so called spacetime-predictor $\stpredictor (\bm{x},t)$ into the weak form and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\intdcell{
  \testfunction{k} \sbasis{l}
}
\right)
(\bm{u^{n+1} - u^{n}})
&+
\left(\intdt{\intdcell{
      \testfunction{k} \Riemann(\stpredictor^-, \stpredictor^+) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\testfunction{k}} \cdot  \flux(\stpredictor)
}}\right)
&=
\left(\intdt{\intdcell{
      \testfunction{k} \source(\stpredictor)
}}\right)
\end{split}
\end{align}
The first term motivates an orthogonal basis as this leads to a diagonal mass matrix.

\sidetitle{Predictor}
To derive the predictor, we again take an approximation of our solution in a nodal basis, but now consider polynomials that are defined both in space and time
\begin{equation}
  \label{eq:cell-approx-space-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis{l}(\bm{x}, t).
\end{equation}
We now multiply the conservation law again by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t)
    \left(
      \divergence{\flux(\stpredictor)}
    \right)
}}
=
\intdt{\intdcell{
  \testfunction{k} \source(\stpredictor)
}}.
\end{equation}
Similar to the derivation of the corrector, we again integrate the first term by parts in time and the flux divergence in space.
This time we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
Note that this neglects the interaction with neighbouring cells; this is corrected in the corrector step.
\todo{Replace $\bm{x}$ with a macro some other letter. Is also coordinate!}

\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \testfunction{k} (\bm{x}, t^{n+1}) \stpredictor(\bm{x}. t^{n+1})
}
&-
\intdt{\intdcell{
    \pdv{}{d} \testfunction{k}(\bm{x}, t) \stpredictor(\bm{x}, t)
}}
-
\intdcell{
  \testfunction{k}(\bm{x},t^n) \dsol(\bm{x}, t^n)
} = \\
\intdt{\intdcell{
    \testfunction{k}(\bm{x},t) \divergence{\flux(\stpredictor)}
}}
&+
\intdt{\intdcell{
    \testfunction{k}(\bm{x}, t) \source(\stpredictor)
}}
\end{split}
\end{align}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be solved in a fixed point iteration scheme.
\todo{Inital guess correct? Maybe cite sth?}
As initial value for the space-time-predictor we use the solution of the timestep before.
\todo{Maybe collect into matrices and show scheme?}
For details and proof of convergence for the linear case, see~\cite{dumbser2008unified}.

\sidetitle{Boundary extrapolation and stuff}
\todo{Explain that stpredictor and fluxes are saved and extrapolated to boundary.
This is important because we extrapolate the gradient as well!}
Extrapolate unknowns to boundary, convert from our basis to basis on faces.
Then Riemann solver
and so on

@@

\sidetitle{Riemann solver \textit{\&} timestep}
As a Riemann solver we use a simple Rusanov-flux that is adapted for diffusive problems
\begin{equation}
  \label{eq:rusanov-flux}
  \Riemann(q_h^-, \nabla q_h^-; g_h^+, \nabla q_h^+) \cdot \bm{n} =
  \frac{1}{2} \left(
    F(q_h^+, \nabla q_h^+) +
    F(q_h^-, \nabla q_h^-)
  \right) -
  \frac{1}{2} s_\text{max} (q_h^+ - q_h^-),
\end{equation}
with a penalty term
\begin{equation}
  \label{eq:parabolic-penalty}
  s_\text{max}  = \max \left(
\maxConvEigen[q_h^-], \, \maxConvEigen[q_h^+]
\right) +
2 \eta \max \left(
\maxViscEigen[q_h^-], \, \maxViscEigen[q_h^+]
\right)
\end{equation}
and
\begin{equation}
  \eta = \frac{2N+1}{h \sqrt{\frac{1}{2} \pi}}.
\end{equation}
In this equation, $N$ is the polynomial order and $h$ is the side-length of an element.
The penalty term depends on the maximal absolute eigenvalues $\lambda_c$ and $\lambda_v$ \todo{
Eigenvalues of jacobian (in normal direction), maximal eigenvalues.}.
Max eigenvalue of convective part $\maxConvEigen$,i.e.\ of $\left( \partial \bm{F}/\partial \bm{Q}\right) \cdot \bm{n}$,
and viscous part $\maxViscEigen$, i.e.\ of $\left( \partial \bm{F}/\partial \left( \nabla \bm{Q} \cdot \bm{n} \right)\right) \cdot \bm{n}$.

This Riemann solver was first published in~\cite{gassner2008discontinuous} and used for an \textsc{ader-dg} scheme in~\cite{dumbser2010arbitrary}.
The timestep is restriced to a so-called \textsc{cfl}-type penalty
\begin{equation}
 \Delta t \leq  \text{CFL} \, \frac{\alpha(N) \, h}{\maxConvEigen + 2 \maxViscEigen \frac{2N+1}{h}}
\end{equation}
with $N$ polynomial order and $h$ characteristic length scale of elements~\cite{dumbser2010arbitrary,gassner2008discontinuous}.
The constant $\alpha(N) \leq {\left( 2N+1  \right)}^{-1}$ is obtained from von Neumann analysis on a simple model problem and depends on the approximation order~\cite{dumbser2008unified}.
We use a value of $0.7$ for the constant $\text{CFL}$.
\todo{Describe CFL-cond.\ for multiple dimensions!}

\section{Reactive Compressible Navier Stokes Equations}
Fluid motion can be described by the compressible Navier Stokes equations.
We follow the description in~\cite{dumbser2010arbitrary} for the Navier Stokes part.
The coupling with the advection-diffusion-reaction equation follows~\cite{hidalgo2011ader} which describes this equation set for the one-dimensional case.

The vector of conserved quantities is given by
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
\end{equation}
where $\Qrho$ is the density, $\Qj$ is the momentum, $\QE$ the energy density and $\QZ$ is the mass fraction of the chemical reactant.

We split the flux into a hyperbolic part $\hyperFlux(\Q)$ and an viscous flux $\viscFlux(\gradQ)$
\begin{equation}
  \label{eq:flux}
  \flux(\Q, \gradQ) = \hyperFlux(\Q) + \viscFlux(\Q, \gradQ).
\end{equation}

@@

Complete \textsc{pde}:
\begin{equation}
 \begin{array}{l}
 \text{mass cons.} \\
 \text{momentum cons.} \\
 \text{energy cons.} \\
 \text{cont.\ gas} 
\end{array}
:
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE\\
    \QZ
    \end{pmatrix}}_{\Q}
  + 
  \divergence{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}
 =
  \underbrace{
  \begin{pmatrix}
    S_{\Qrho\phantom{\Qrho}}\\
    S_{\Qj}\\
    S_{\QE}\\
    S_{\QZ}
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}

@@

A relatively recent way of dealing with this problem is the finite volume subcell limiter.
Our discussion follows the one in~\cite{dumbser2016simple}.
This limiting is an a posteriori limiting, which means that we first evaluate a timestep with an unlimited \textsc{ader-dg} method and then check, whether our solution is \enquote{correct}.
In our case, we check whether the solution is a valid floating point number.
We also check whether it is physically admissible.
In our case we use the criterion
\begin{equation}
  \label{eq:limiting-physical}
  \operatorname{is-admissible}(\Q) =
  \begin{cases}
    \text{true} & \Qrho > 0 \land \pressure > 0 \land \QZ > 0\\
    \text{false} & \text{otherwise}.
  \end{cases}
\end{equation}
The positivity of the pressure implies positivity of energy.
\todo{Mention dmp if I end up using it.}
